### 2013-12-11 manuscripts/dpp-msbayes/dpp-msbayes-si-figures.tex
Working on dpp ms.
Incorporating new plots into SI figures (`dpp-msbayes-si-figures.tex`).
Adding some macros to `macros.tex` for the new figure captions.
Also, adding SI to the main `dpp-msbayes.tex` file.

%% Validation accuracy
\siFigure{../../validation/images/validation-accuracy-psi.pdf}{
    \validationAccuracyComparisonCaption{unadjusted}{\divTimeNum}
    \jitterComment
}{figValidationAccuracyPsi}

\siFigure{../../validation/images/validation-accuracy-psi-glm.pdf}{
    \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeNum}
    \jitterComment
}{figValidationAccuracyPsiGlm}

\siFigure{../../validation/images/validation-accuracy-omega.pdf}{
    \validationAccuracyComparisonCaption{unadjusted}{\divTimeDispersion}
}{figValidationAccuracyOmega}

\siFigure{../../validation/images/validation-accuracy-omega-glm.pdf}{
    \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeDispersion}
}{figValidationAccuracyOmegaGlm}

\siFigure{../../validation/images/validation-accuracy-time.pdf}{
    \validationAccuracyComparisonCaption{unadjusted}{\divTimeMean}
}{figValidationAccuracyTime}

\siFigure{../../validation/images/validation-accuracy-time-glm.pdf}{
    \validationAccuracyComparisonCaption{GLM-adjusted}{\divTimeMean}
}{figValidationAccuracyTimeGlm}

%% Validation model choice
\siFigure{../../validation/images/validation-model-choice-psi.pdf}{
    \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeNum = 1$}
}{figValidationModelChoicePsi}

\siFigure{../../validation/images/validation-model-choice-psi-glm.pdf}{
    \validationModelChoiceComparisonCaption{GLM-adjusted}{$\divTimeNum = 1$}
}{figValidationModelChoicePsiGlm}

\siFigure{../../validation/images/validation-model-choice-omega.pdf}{
    \validationModelChoiceComparisonCaption{unadjusted}{$\divTimeDispersion < 0.01$}
}{figValidationModelChoiceOmega}

\siFigure{../../validation/images/validation-model-choice-omega-glm.pdf}{
    \validationModelChoiceComparisonCaption{GLM-adjusted}{$\divTimeDispersion < 0.01$}
}{figValidationModelChoiceOmegaGlm}

%% Validation ordered models
\siFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_accuracy.pdf}{
    \validationAccuracyCaption{\modelDPPOrdered}{\modelDPPOrdered}
}{figValidationAccuracyDPPOrderedDPPOrdered}

\siFigure{../../validation/no-sort/results/pymsbayes-results/plots/prior-dpp_prior-dpp_mc_behavior.pdf}{
    \validationModelChoiceCaption{\modelDPPOrdered}{\modelDPPOrdered}
}{figValidationModelChoiceDPPOrderedDPPOrdered}

%% Power accuracy results
\siFigure{../../power-comparison/images/old-power-omega-accuracy-6.pdf}{
    \powerAccuracyComparisonCaption{\powerSeriesOld}
}{figPowerAccuracyOld}

\siFigure{../../power-comparison/images/uniform-power-omega-accuracy-6.pdf}{
    \powerAccuracyComparisonCaption{\powerSeriesUniform}
}{figPowerAccuracyUniform}

\siFigure{../../power-comparison/images/exp-power-omega-accuracy-6.pdf}{
    \powerAccuracyComparisonCaption{\powerSeriesExp}
}{figPowerAccuracyExp}

%% Power psi results
\siFigure{../../power-comparison/images/old-power-psi-6.pdf}{
    \powerComment{\powerSeriesOld}
    \powerPsiComment{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siFigure{../../power-comparison/images/uniform-power-psi-6.pdf}{
    \powerComment{\powerSeriesUniform}
    \powerPsiComment{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siFigure{../../power-comparison/images/exp-power-psi-6.pdf}{
    \powerComment{\powerSeriesExp}
    \powerPsiComment{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}

%% Power dispersion results
\siFigure{../../power-comparison/images/old-power-omega-6.pdf}{
    \powerComment{\powerSeriesOld}
    \powerDispersionComment{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siFigure{../../power-comparison/images/uniform-power-omega-6.pdf}{
    \powerComment{\powerSeriesUniform}
    \powerDispersionComment{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siFigure{../../power-comparison/images/exp-power-omega-6.pdf}{
    \powerComment{\powerSeriesExp}
    \powerDispersionComment{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}

%% Power psi probability results
\siFigure{../../power-comparison/images/old-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesOld}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siFigure{../../power-comparison/images/uniform-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesUniform}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siFigure{../../power-comparison/images/exp-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesExp}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}

%% Power dispersion prob results
\siFigure{../../power-comparison/images/old-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesOld}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siFigure{../../power-comparison/images/uniform-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesUniform}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siFigure{../../power-comparison/images/exp-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesExp}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}