### 2013-12-26 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex

%:TABLE-notation descriptions
\begin{table}[htbp]
    \sffamily
    \scriptsize
    \rowcolors{2}{}{myGray}
    \addtolength{\tabcolsep}{-0.1cm}
\caption{Summary of the notation used throughout this work; modified from \citet{Oaks2012}.}
    \centering
    \begin{tabular}{ l >{\raggedright\hangindent=0.5cm}m{14cm} }
        \toprule
        \textbf{Symbol} & \textbf{Description} \tn
        \midrule
        \npairs{} & Number of population pairs. \tn
        \popSampleSize{i}{} & The number of genome copies sampled from population pair $i$, with \popSampleSize{1}{i} sampled from population 1, and \popSampleSize{2}{i} from population 2. \tn
        \nloci{i} & Number of loci sampled from population pair $i$. \tn
        \nlociTotal & Total number of unique loci sampled. \tn
        \alignment{i}{j} & Sequence alignment of locus $j$ sampled from population pair $i$. \tn
        \alignmentSSObs{i}{j} & Population genetic summary statistics calculated from \alignment{i}{j}. \tn
        \alignmentVector & Vector containing the sequence alignments of each locus from each population pair: $(\alignment{1}{1},\ldots,\alignment{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \ssVectorObs & Vector containing the summary statistics of each locus from each population pair: $(\alignmentSSObs{1}{1},\ldots,\alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \ssSpace & Multi-dimensional Euclidean space around the observed summary statistics, \ssVectorObs. \tn
        \tol & Radius of \ssSpace, i.e., the tolerance of the ABC estimation. \tn
        \geneTree{i}{j} & Gene tree of the sequences in \alignment{i}{j}. \tn
        \geneTreeVector & Vector containing the gene trees of each locus from each population pair: $(\geneTree{1}{1},\ldots,\geneTree{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \divTimeNum & Number of unique population divergence times among the \npairs{} population pairs. \tn
        \divTime{} & One of the \divTimeNum \iid divergence time paramters in \globalcoalunit generations. \tn
        \divTimeVector & Ordered set of divergence time parameters: $\{\divTime{1},\ldots,\divTime{\divTimeNum}\}$. \tn
        \divTimeIndex{i} & The index of the divergence time parameter in \divTimeVector that is mapped to population pair $i$. \tn
        \divTimeIndexVector & Vector of divergence time parameter indices: $(\divTimeIndex{1},\ldots,\divTimeIndex{\npairs{}})$. \tn
        \divTimeMap{i} & Time of divergence in \globalcoalunit generations between the populations of pair $i$. \tn
        \divTimeMapVector & Vector of divergence times for each of the population pairs: $(\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})$. \tn
        \divTimeScaled{i}{j} & Scaled time of divergence between the populations of pair $i$ for locus $j$. \tn
        \divTimeScaledVector & Vector containing the scaled divergence times of each locus from each population pair: $(\divTimeScaled{1}{1},\ldots,\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \descendantTheta{1}{i}, \descendantTheta{2}{i} & \myTheta{} of the $1^{st}$ and $2^{nd}$ descendent population, respectively, of pair $i$. \tn
        \ancestralTheta{i} & \myTheta{} of the population ancestral to pair $i$. \tn
        \descendantThetaVector{1}, \descendantThetaVector{2} & Vectors $(\descendantTheta{1}{1},\ldots,\descendantTheta{1}{\npairs{}})$ and $(\descendantTheta{2}{1},\ldots,\descendantTheta{2}{\npairs{}})$, respectively. \tn
        \ancestralThetaVector & Vector containing the \ancestralTheta{} parameters for each population pair: $(\ancestralTheta{1},\ldots,\ancestralTheta{\npairs{}})$. \tn
        \locusMutationRateScalar{j} & mutation-rate scaling parameter for locus $j$. \tn
        \locusMutationRateScalarVector & Vector containing the locus-specific mutation-rate parameters: $(\locusMutationRateScalar{1},\ldots,\locusMutationRateScalar{\nlociTotal})$. \tn
        \locusRateHetShapeParameter & Hyper-parameter for the shape of the gamma prior distribution on each \locusMutationRateScalar{}. \tn
        \bottleScalar{1}{i}, \bottleScalar{2}{i} & \myTheta{}-scaling parameters that determine the magnitude of the population bottleneck in the $1^{st}$ and $2^{nd}$ descendant population of pair $i$, respectively. The bottleneck in each descendant population begins immediately after divergence. \tn
        \bottleScalarVector{1}, \bottleScalarVector{2} & Vectors $(\bottleScalar{1}{1},\ldots,\bottleScalar{1}{\npairs{}})$ and $(\bottleScalar{2}{1},\ldots,\bottleScalar{2}{\npairs{}})$, respectively. \tn
        \bottleTime{i} & Proportion of time between present and \divt{i} when the bottleneck ends for the descendant populations of pair $i$. \tn
        \bottleTimeVector & Vector containing the \bottleTime{} parameters for each population pair: $(\bottleTime{1},\ldots,\bottleTime{\npairs{}})$. \tn
        \migrationRate{i} & Symmetric migration rate between the descendant populations of pair $i$. \tn
        \migrationRateVector & Vector containing the migration rates for each population pair: $(\migrationRate{1},\ldots,\migrationRate{\npairs{}})$. \tn
        \ploidyScalar{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$. This constant is intended to allow the user to scale \myTheta{} for differences in ploidy among loci or differences in generation times among taxa. \tn
        \mutationRateScalarConstant{i}{j} & \myTheta{}-scaling constant provided by the user for locus $j$ of pair $i$. This constant is intended to allow the user to scale \myTheta{} for differences in mutation rates among loci or among taxa. \tn
        \ploidyScalarVector & Vector of ploidy and/or generation-time scaling constants: $(\ploidyScalar{1}{1},\ldots,\ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \mutationRateScalarConstantVector & Vector of mutation-rate scaling constants: $(\mutationRateScalarConstant{1}{1},\ldots,\mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \divTimeMean  & Mean of \divTimeMapVector. \tn
        \divTimeVar  & Variance of \divTimeMapVector. \tn
        \divTimeDispersion  & Dispersion index of \divTimeMapVector ($\divTimeVar/\divTimeMean$). \tn
        \numPriorSamples & Number of samples from the joint prior. \tn
        \paramSampleVector{} & Vector of parameter values drawn from the joint prior. \tn
        \ssVector{} & Vector containing the summary statistics calculated from data simulated under parameter values drawn from the prior (\hpvector{}). \tn
        \paramSampleMatrix{} & Random sample of $\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples}$ drawn form the prior. \tn
        \ssMatrix & Summary statistic vectors $\ssVector{1}, \ldots, \ssVector{\numPriorSamples}$ for each $\hpvector{1}, \ldots, \hpvector{\numPriorSamples}$ drawn from the prior. \tn
        \bottomrule
    \end{tabular}
    \label{tabNotation}
\end{table}

\clearpage

\begin{table}[htbp]
    \sffamily
    % \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{The model priors evaluated in our simulation-based analyses.  For
        the \modelDPP model, the prior on the concentration parameter,
        $\concentrationParam \sim Gamma(\cdot,\cdot)$, was set to $Gamma(2,2)$
        for the validation analyses and $Gamma(1.5,18.1)$ for the power
        analyses.}
    \centering
    \begin{tabular}{ l l l l l }
        \toprule
        & \multicolumn{4}{c}{Priors} \\
        \cmidrule(){2-5}
        Model & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
        \midrule
            \modelOld & $\divTimeIndexVector \sim \priorOld$
                      & $\divTime{} \sim U(0,10)$
                      & $\ancestralTheta{} \sim U(0, 0.05)$
                      & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(\cdot,\cdot)}$
                      & $\divTime{} \sim Exp(0.3464)$
                      & $\ancestralTheta{} \sim Exp(40)$
                      & $\descendantTheta{}{} \sim Exp(40)$ \\
            \modelUniform & $\divTimeIndexVector \sim \priorUniform$
                          & $\divTime{} \sim Exp(0.3464)$
                          & $\ancestralTheta{} \sim Exp(40)$
                          & $\descendantTheta{}{} \sim Exp(40)$ \\
            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
                          & $\divTime{} \sim Exp(0.3464)$
                          & $\ancestralTheta{} \sim Exp(40)$
                          & $\descendantTheta{}{} \sim Exp(40)$ \\
        \bottomrule
    \end{tabular}
    \label{tabPriors}
\end{table}

\clearpage

\begin{table}[htbp]
    \sffamily
    \footnotesize
    % \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{The models we used to simulate pseudo-replicate datasets for
        assessing the power of the models in Table \ref{tabPriors}.
        The distributions of divergence times are given in units of
        \globalcoalunit generations and millions of generations ago (MGA), the
        latter assuming a per-site rate of 1\e{-8} mutations per generation.}
    \centering
    \begin{tabular}{ l l l l l l }
        \toprule
        & \multicolumn{5}{c}{Priors} \\
        \cmidrule(){2-6}
        Model series & \divTimeIndexVector & \divTime{} (\globalcoalunit gens.) & \divTime{} (MGA) & \ancestralTheta{} & \descendantTheta{}{} \\
        \midrule
            \powerSeriesOld & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.2)$
                            & $\divTime{} \sim U(0,0.5)$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.4)$
                            & $\divTime{} \sim U(0,1.0)$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.6)$
                            & $\divTime{} \sim U(0,1.5)$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.8)$
                            & $\divTime{} \sim U(0,2.0)$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,1.0)$
                            & $\divTime{} \sim U(0,2.5)$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,2.0)$
                            & $\divTime{} \sim U(0,5.0)$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
        \midrule
            \powerSeriesUniform & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.2)$
                            & $\divTime{} \sim U(0,0.5)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.4)$
                            & $\divTime{} \sim U(0,1.0)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.6)$
                            & $\divTime{} \sim U(0,1.5)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.8)$
                            & $\divTime{} \sim U(0,2.0)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,1.0)$
                            & $\divTime{} \sim U(0,2.5)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,2.0)$
                            & $\divTime{} \sim U(0,5.0)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
        \midrule
            \powerSeriesExp & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(17.32)$
                            & $\divTime{} \sim Exp(mean = 0.14)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(8.66)$
                            & $\divTime{} \sim Exp(mean = 0.29)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(5.77)$
                            & $\divTime{} \sim Exp(mean = 0.43)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(4.33)$
                            & $\divTime{} \sim Exp(mean = 0.58)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(3.46)$
                            & $\divTime{} \sim Exp(mean = 0.72)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(1.73)$
                            & $\divTime{} \sim Exp(mean = 1.44)$
                            & $\ancestralTheta{} \sim Exp(40)$
                            & $\descendantTheta{}{} \sim Exp(40)$ \\
        \bottomrule
    \end{tabular}
    \label{tabPowerModels}
\end{table}

\begin{table}[htbp]
    \sffamily
    \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{The models used to analyze the data from the 22 pairs of taxa from
        the Philippines ($\mathbf{M}$), and a subset of nine of those pairs
        from the Islands of Negros and Panay ($\mathbb{M}$).
        In addition to the $\popSampleSize{}{} - 1$ coalescent times, the
        \empModelDPPSimple has only a single \myTheta{} parameter for
        each taxon pair.
        The remaining $\mathbf{M}$ models have three \myTheta{}, two
        \bottleScalar{}{}, and one \bottleTime{} parameters.
        The distributions of divergence times are given in units of
        \globalcoalunit generations and millions of generations ago (MGA), the
        latter assuming a per-site rate of 1\e{-8} mutations per generation.
        The \npModelDPP model (and its \npModelDPPOrdered counterpart
        that samples over ordered divergence models) has only two
        \myTheta{} parameters (the descendant populations of each
        pair share the same \myTheta{} parameter, and there are no
        bottleneck parameters).}
    \centering
    \begin{tabular}{ l l l l l l l }
        \toprule
        & \multicolumn{6}{c}{Priors} \\
        \cmidrule(){2-7}
        Model & \divTimeIndexVector & \divTime{} (\globalcoalunit gens.) & \divTime{} (MGA) & \ancestralTheta{} & \descendantTheta{}{} & \bottleScalar{}{} \\
        \midrule
            \empModelOld & $\divTimeIndexVector \sim \priorOld$
                      & $\divTime{} \sim U(0,34.64)$
                      & $\divTime{} \sim U(0,17.3)$
                      & $\ancestralTheta{} \sim U(0, 0.01)$
                      & $\descendantThetaMean{} \sim U(0, 0.01)$ 
                      & $\bottleScalar{}{} \sim U(0, 1)$ \\
            \empModelUniform & $\divTimeIndexVector \sim \priorUniform$
                      & $\divTime{} \sim Exp(0.1)$
                      & $\divTime{} \sim Exp(mean=5)$
                      & $\ancestralTheta{} \sim Exp(200)$
                      & $\descendantTheta{}{} \sim Exp(200)$
                      & $\bottleScalar{}{} \sim Beta(5, 1)$ \\
            \empModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
                      & $\divTime{} \sim Exp(0.1)$
                      & $\divTime{} \sim Exp(mean=5)$
                      & $\ancestralTheta{} \sim Exp(200)$
                      & $\descendantTheta{}{} \sim Exp(200)$
                      & $\bottleScalar{}{} \sim Beta(5, 1)$ \\
            \empModelDPPInform & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
                      & $\divTime{} \sim Exp(1/6)$
                      & $\divTime{} \sim Exp(mean=3)$
                      & $\ancestralTheta{} \sim Exp(200)$
                      & $\descendantTheta{}{} \sim Exp(200)$
                      & $\bottleScalar{}{} \sim Beta(5, 1)$ \\
            \empModelDPPSimple & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
                      & $\divTime{} \sim Exp(0.1)$
                      & $\divTime{} \sim Exp(mean=5)$
                      & $\myTheta{} \sim Exp(200)$
                      & -
                      & - \\
            \npModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,5.0)}$
                      & $\divTime{} \sim Exp(0.1)$
                      & $\divTime{} \sim Exp(mean=5)$
                      & $\ancestralTheta{} \sim Exp(200)$
                      & $\descendantTheta{}{} \sim Exp(200)$
                      & - \\
        \bottomrule
    \end{tabular}
    \label{tabEmpiricalModels}
\end{table}