### 2014-02-22 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex

%:TABLE-notation descriptions
\begin{table}[htbp]
    \sffamily
    \scriptsize
    \rowcolors{2}{}{myGray}
    \addtolength{\tabcolsep}{-0.1cm}
\caption{Summary of the notation used throughout this work; modified from \citet{Oaks2012}.}
    \centering
    \begin{tabular}{ l >{\raggedright\hangindent=0.5cm}m{14cm} }
        \toprule
        \textbf{Symbol} & \textbf{Description} \tn
        \midrule
        \npairs{} & Number of population pairs. \tn
        \popSampleSize{i}{} & The number of genome copies sampled from population pair $i$, with \popSampleSize{1}{i} sampled from population 1, and \popSampleSize{2}{i} from population 2. \tn
        \nloci{i} & Number of loci sampled from population pair $i$. \tn
        \nlociTotal & Total number of unique loci sampled. \tn
        \alignment{i}{j} & Sequence alignment of locus $j$ sampled from population pair $i$. \tn
        \alignmentSSObs{i}{j} & Population genetic summary statistics calculated from \alignment{i}{j}. \tn
        \alignmentVector & Vector containing the sequence alignments of each locus from each population pair: $(\alignment{1}{1},\ldots,\alignment{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \ssVectorObs & Vector containing the summary statistics of each locus from each population pair: $(\alignmentSSObs{1}{1},\ldots,\alignmentSSObs{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \ssSpace & Multi-dimensional Euclidean space around the observed summary statistics, \ssVectorObs. \tn
        \tol & Radius of \ssSpace, i.e., the tolerance of the ABC estimation. \tn
        \geneTree{i}{j} & Gene tree of the sequences in \alignment{i}{j}. \tn
        \geneTreeVector & Vector containing the gene trees of each locus from each population pair: $(\geneTree{1}{1},\ldots,\geneTree{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \divTimeNum & Number of population divergence-time parameters shared among the \npairs{} population pairs. \tn
        \divTime{} & Time of population divergence in \globalcoalunit generations. \tn
        \divTimeVector & Set of divergence-time parameters: $\{\divTime{1},\ldots,\divTime{\divTimeNum}\}$. \tn
        \divTimeIndex{i} & The index of the divergence-time in \divTimeVector to which population pair $i$ is mapped. \tn
        \divTimeIndexVector & Vector of divergence-time indices: $(\divTimeIndex{1},\ldots,\divTimeIndex{\npairs{}})$. \tn
        \divTimeMap{i} & Time of divergence in \globalcoalunit generations between the populations of pair $i$. \tn
        \divTimeMapVector & Vector of divergence times for each of the population pairs: $(\divTimeMap{1},\ldots,\divTimeMap{\npairs{}})$. \tn
        \divTimeScaled{i}{j} & Scaled time of divergence between the populations of pair $i$ for locus $j$. \tn
        \divTimeScaledVector & Vector containing the scaled divergence times of each locus from each population pair: $(\divTimeScaled{1}{1},\ldots,\divTimeScaled{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \descendantTheta{1}{i}, \descendantTheta{2}{i} & Mutation-rate-scaled effective population size of the $1^{st}$ and $2^{nd}$ descendent population, respectively, of pair $i$. \tn
        \ancestralTheta{i} & Mutation-rate-scaled effective population size of the population ancestral to pair $i$. \tn
        \descendantThetaVector{1}, \descendantThetaVector{2} & Vectors $(\descendantTheta{1}{1},\ldots,\descendantTheta{1}{\npairs{}})$ and $(\descendantTheta{2}{1},\ldots,\descendantTheta{2}{\npairs{}})$, respectively. \tn
        \ancestralThetaVector & Vector containing the \ancestralTheta{} parameters for each population pair: $(\ancestralTheta{1},\ldots,\ancestralTheta{\npairs{}})$. \tn
        \locusMutationRateScalar{j} & Mutation-rate multiplier of locus $j$. \tn
        \locusMutationRateScalarVector & Vector containing the locus-specific mutation-rate multipliers: $(\locusMutationRateScalar{1},\ldots,\locusMutationRateScalar{\nlociTotal})$. \tn
        \locusRateHetShapeParameter & The shape parameter of the gamma prior distribution on \locusMutationRateScalar{}. \tn
        \bottleScalar{1}{i}, \bottleScalar{2}{i} & \myTheta{}-scaling parameters that determine the magnitude of the population bottleneck in the $1^{st}$ and $2^{nd}$ descendant population of pair $i$, respectively. The bottleneck in each descendant population begins immediately after divergence. \tn
        \bottleScalarVector{1}, \bottleScalarVector{2} & Vectors $(\bottleScalar{1}{1},\ldots,\bottleScalar{1}{\npairs{}})$ and $(\bottleScalar{2}{1},\ldots,\bottleScalar{2}{\npairs{}})$, respectively. \tn
        \bottleTime{i} & Proportion of time between present and \divTimeMap{i} when the bottleneck ends for the descendant populations of pair $i$. \tn
        \bottleTimeVector & Vector containing the \bottleTime{} parameters for each population pair: $(\bottleTime{1},\ldots,\bottleTime{\npairs{}})$. \tn
        \migrationRate{i} & Symmetric migration rate between the descendant populations of pair $i$. \tn
        \migrationRateVector & Vector containing the migration rates for each population pair: $(\migrationRate{1},\ldots,\migrationRate{\npairs{}})$. \tn
        \ploidyScalar{i}{j} & \myTheta{}-scaling constant provided by the investigator for locus $j$ of pair $i$. This constant is required to scale \myTheta{} for differences in ploidy among loci or differences in generation times among taxa. \tn
        \mutationRateScalarConstant{i}{j} & \myTheta{}-scaling constant provided by the investigator for locus $j$ of pair $i$. This constant is required to scale \myTheta{} for differences in mutation rates among loci or among taxa. \tn
        \ploidyScalarVector & Vector of ploidy and/or generation-time scaling constants: $(\ploidyScalar{1}{1},\ldots,\ploidyScalar{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \mutationRateScalarConstantVector & Vector of mutation-rate scaling constants: $(\mutationRateScalarConstant{1}{1},\ldots,\mutationRateScalarConstant{\npairs{}}{\nloci{\npairs{}}})$. \tn
        \divTimeMean  & Mean of divergence times across the \npairs{} population pairs. \tn
        \divTimeVar  & Variance of divergence times across the \npairs{} population pairs. \tn
        \divTimeDispersion  & Dispersion index of divergence times across the \npairs{} population pairs ($\divTimeVar/\divTimeMean$). \tn
        \numPriorSamples & Number of samples from the joint prior. \tn
        \paramSampleVector{} & Vector of parameter values drawn from the joint prior. \tn
        \ssVector{} & Vector containing the summary statistics calculated from data simulated under parameter values drawn from the prior (\hpvector{}). \tn
        \paramSampleMatrix{} & Random sample of $\paramSampleVector{1}, \ldots, \paramSampleVector{\numPriorSamples}$ drawn form the prior. \tn
        \ssMatrix & Summary statistic vectors $\ssVector{1}, \ldots, \ssVector{\numPriorSamples}$ for each $\hpvector{1}, \ldots, \hpvector{\numPriorSamples}$ drawn from the prior. \tn
        \bottomrule
    \end{tabular}
    \label{tabNotation}
\end{table}

\clearpage

\begin{table}[htbp]
    \sffamily
    \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{The models evaluated in the simulation-based analyses.  For
        the \modelDPP model, the prior on the concentration parameter,
        $\concentrationParam \sim Gamma(\cdot,\cdot)$, was set to $Gamma(2,2)$
        for the validation analyses and $Gamma(1.5,18.1)$ for the power
        analyses. The distributions of divergence times are given in units of
        \globalcoalunit generations followed in brackets by units of millions
        of generations ago (MGA), with the former converted to the latter
        assuming a per-site rate of 1\e{-8} mutations per generation. For model
        \modelOld, the priors for theta parameters are $\ancestralTheta{} \sim
        U(0, 0.05)$ and $\descendantTheta{1}{}, \descendantTheta{2}{} \sim
        Beta(1, 1) \times 2 \times U(0, 0.05)$. The later is summarized as
        $\descendantThetaMean{} \sim U(0, 0.05)$. For the \modelDPP and
        \modelUniform, and \modelUshaped models, \ancestralTheta{},
        \descendantTheta{1}{}, and \descendantTheta{2}{} are independently
        and exponentially distributed with a mean of 0.025.}
    \centering
    \begin{tabular}{ l l l l l }
        \toprule
        & \multicolumn{4}{c}{Priors} \\
        \cmidrule(){2-5}
        Model & \divTimeIndexVector & \divTime{} & \myTheta{} &  \\
        \midrule
            \modelOld & $\divTimeIndexVector \sim \priorOld$
                      & $\divTime{} \sim U(0,10 \; [25 \; MGA])$
                      & $\ancestralTheta{} \sim U(0, 0.05)$
                      & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
                          & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
                          & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
            \modelUniform & $\divTimeIndexVector \sim \priorUniform$
                          & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
                          & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(\cdot,\cdot)}$
                      & $\divTime{} \sim Exp(mean=2.887 \; [7.22 \; MGA])$
                      & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
        \bottomrule
    \end{tabular}
    \label{tabPriors}
\end{table}

\clearpage

\begin{table}[htbp]
    \sffamily
    \footnotesize
    % \scriptsize
    % \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    % \captionsetup{font=footnotesize}
    \caption{The models used to simulate pseudo-replicate datasets for
        assessing the power of the models in Table \ref{tabPriors}.  The
        distributions of divergence times are given in units of \globalcoalunit
        generations followed in brackets by units of millions of generations
        ago (MGA), with the former converted to the latter assuming a per-site
        rate of 1\e{-8} mutations per generation. For all of the
        \powerSeriesOld models, the priors for theta parameters are
        $\ancestralTheta{} \sim U(0, 0.05)$ and $\descendantTheta{1}{},
        \descendantTheta{2}{} \sim Beta(1, 1) \times 2 \times U(0, 0.05)$. The
        later is summarized as $\descendantThetaMean{} \sim U(0, 0.05)$. For
        the \powerSeriesUniform and \powerSeriesExp models, \ancestralTheta{},
        \descendantTheta{1}{}, and \descendantTheta{2}{} are independently
        and exponentially distributed with a mean of 0.025.}
    \centering
    \begin{tabular}{ l l l l l }
        \toprule
        & \multicolumn{4}{c}{Priors} \\
        \cmidrule(){2-5}
        Model series & \divTimeIndexVector & \divTime{} & \myTheta{} &  \\
        \midrule
            \powerSeriesOld & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.2 \; [0.5 \; MGA])$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.4 \; [1.0 \; MGA])$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.6 \; [1.5 \; MGA])$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.8 \; [2.0 \; MGA])$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,1.0 \; [2.5 \; MGA])$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,2.0 \; [5.0 \; MGA])$
                            & $\ancestralTheta{} \sim U(0, 0.05)$
                            & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
        \midrule
            \powerSeriesUniform & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.2 \; [0.5 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.4 \; [1.0 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.6 \; [1.5 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,0.8 \; [2.0 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,1.0 \; [2.5 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim U(0,2.0 \; [5.0 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
        \midrule
            \powerSeriesExp & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(mean = 0.058 \; [0.14 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(mean = 0.115 \; [0.29 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(mean = 0.173 \; [0.43 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(mean = 0.231 \; [0.58 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(mean = 0.289 \; [0.72 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
                            & $\divTimeNum = 22$
                            & $\divTime{} \sim Exp(mean = 0.577 \; [1.44 \; MGA])$
                            & \multicolumn{2}{l}{$\ancestralTheta{} \sim \descendantTheta{1}{} \sim \descendantTheta{2}{} \sim Exp(mean=0.025)$} \\
        \bottomrule
    \end{tabular}
    \label{tabPowerModels}
\end{table}

\begin{table}[htbp]
    \sffamily
    \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{myGray}{}
    %\captionsetup{font=footnotesize}
    \caption{The models used to analyze the data from the 22 pairs of taxa from
        the Philippines ($\mathbf{M}$), and a subset of nine of those pairs
        from the Islands of Negros and Panay ($\mathbb{M}$).
        In addition to the $\popSampleSize{}{} - 1$ coalescent times, the
        \empModelDPPSimple has only a single \myTheta{} parameter for
        each taxon pair.
        The remaining $\mathbf{M}$ models have three \myTheta{}, two
        \bottleScalar{}{}, and one \bottleTime{} parameter.
        The distributions of divergence times are given in units of
        \globalcoalunit generations followed in brackets by units of millions
        of generations ago (MGA), with the former converted to the latter
        assuming a per-site rate of 1\e{-8} mutations per generation.
        The \npModelDPP model (and its \npModelDPPOrdered counterpart
        that samples over ordered divergence models) has only two
        \myTheta{} parameters (the descendant populations of each
        pair share the same \myTheta{} parameter, and there are no
        bottleneck parameters).}
    \centering
    \begin{tabulary}{\textwidth}{ l L }
        \toprule
        Model & Priors \\
        \midrule
            \empModelOld & $\divTimeIndexVector \sim \priorOld$ \tb
        $\divTime{} \sim U(0,34.64 \, [17.3 \; MGA])$ \tb
                      $\ancestralTheta{} \sim U(0, 0.01)$ \tb
                      $\descendantTheta{1}{}, \descendantTheta{2}{} \sim
                            Beta(1, 1) \times 2 \times U(0, 0.01)$ \tb
                      $\bottleScalar{1}{} \sim U(0, 1)$ \tb
                      $\bottleScalar{2}{} \sim U(0, 1)$ \\[0.25em]
            \empModelUniform & $\divTimeIndexVector \sim \priorUniform$ \tb
                      $\divTime{} \sim Exp(mean=10 \; [5 \; MGA])$ \tb
                      $\ancestralTheta{} \sim Exp(mean=0.005)$ \tb
                      $\descendantTheta{1}{} \sim Exp(mean=0.005)$ \tb
                      $\descendantTheta{2}{} \sim Exp(mean=0.005)$ \tb
                      $\bottleScalar{1}{} \sim Beta(5, 1)$ \tb
                      $\bottleScalar{2}{} \sim Beta(5, 1)$ \\[0.25em]
            \empModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$ \tb
                      $\divTime{} \sim Exp(mean=10 \; [5 \; MGA])$ \tb
                      $\ancestralTheta{} \sim Exp(mean=0.005)$ \tb
                      $\descendantTheta{1}{} \sim Exp(mean=0.005)$ \tb
                      $\descendantTheta{2}{} \sim Exp(mean=0.005)$ \tb
                      $\bottleScalar{1}{} \sim Beta(5, 1)$ \tb
                      $\bottleScalar{2}{} \sim Beta(5, 1)$ \tb \\
            \empModelDPPInform & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$ \tb
                      $\divTime{} \sim Exp(mean=6 \; [3 \; MGA])$ \tb
                      $\ancestralTheta{} \sim Exp(mean=0.005)$ \tb
                      $\descendantTheta{1}{} \sim Exp(mean=0.005)$ \tb
                      $\descendantTheta{2}{} \sim Exp(mean=0.005)$ \tb
                      $\bottleScalar{1}{} \sim Beta(5, 1)$ \tb
                      $\bottleScalar{2}{} \sim Beta(5, 1)$ \tb \\
            \empModelDPPSimple & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$ \tb
                      $\divTime{} \sim Exp(mean=10 \; [5 \; MGA])$ \tb
                      $\ancestralTheta{} = \descendantTheta{1}{} = \descendantTheta{2}{} \sim Exp(mean=0.005)$ \tb
                      $\bottleScalar{1}{} = \bottleScalar{2}{} = 1.0$ \\[0.25em]
            \npModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,5.0)}$ \tb
                      $\divTime{} \sim Exp(mean=10 \; [5 \; MGA])$ \tb
                      $\ancestralTheta{} \sim Exp(mean=0.005)$ \tb
                      $\descendantTheta{1}{} = \descendantTheta{2}{} \sim Exp(mean=0.005)$ \tb
                      $\bottleScalar{1}{} = \bottleScalar{2}{} = 1.0$ \\[0.25em]
        \bottomrule
    \end{tabulary}
    \label{tabEmpiricalModels}
\end{table}