### 2014-02-28 manuscripts/reply/cover-letter/cover-letter.tex
Adding cover letter for reply ms.

\documentclass[letterpaper]{letter}
\usepackage{anysize}
\marginsize{1in}{1in}{0.5in}{0.5in}
\usepackage{setspace}
\usepackage{url}
\usepackage{verbatim}
\usepackage{parskip}
\usepackage{hyperref}
\hypersetup{pdfborder={0 0 0}, colorlinks=true, urlcolor=blue}

\newcommand{\ignore}[1]{}
\newcommand{\super}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\sub}[1]{\ensuremath{_{\textrm{#1}}}}
\newcommand{\dC}{\ensuremath{^\circ{\textrm{C}}}}

\makeatletter
\let\@texttop\relax
\makeatother

\newenvironment{myEnumerate}{
  \begin{enumerate}
    \setlength{\itemsep}{1pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}}
  {\end{enumerate}}

\newenvironment{myItemize}{
  \begin{itemize}
    \setlength{\itemsep}{1pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}}
  {\end{itemize}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\signature{Jamie Oaks \\
                 Postdoctoral Fellow \\ 
                 Department of Biology \\
                 University of Washington \\ 
                 Box 351800 \\ 
                 Seattle, WA 98195 \\ 
                 USA \\
                 \href{mailto:joaks1@uw.edu}{\texttt{joaks1@uw.edu}} \\ 
                 \href{http://www.phyletica.com}{\texttt{www.phyletica.com}}}
%\address{Biodiversity Institute \\ University of Kansas \\ Dyche Hall, 1345 Jayhawk Blvd \\ Lawrence, KS 66045}
\begin{document}
\begin{letter}{The Society for the Study of Evolution \\
                     PO Box 299 \\
                     St. Louis, MO 63166-0299}
\opening{Dear Editor:}
% \raggedright
I am submitting a manuscript entitled ``Implications of uniformly distributed,
empirically informed priors for phylogeographical model selection: A reply to
Hickerson et al.'' to be considered for publication in the \emph{International
Journal of Organic Evolution}.
This manuscript is a response to a paper by Hickerson et al.\ (2014,
\emph{Evolution} 68:284--294), which was a response to a paper published last
year by my co-authors and me (Oaks et al.\ 2013, \emph{Evolution}
67:991--1010).
Former Associate Editor Dr.\ Laura Kubatko invited us to write this response,
with the intention of handling its review.
The following is an excerpt from Dr.\ Kubatko's comments in the decision letter
for Hickerson et al.\ (2014):

\begin{quote}
Thus, I propose the following: Hickerson et al. should prepare and
submit the final version of their response.  I urge them to fully
consider all of the points raised by Oaks in his most recent review of their
paper. However, since the point of view of Hickerson et al. seems to
be at odds with Oaks et al., I would like to give Oaks et al. the opportunity
to respond, if they so choose. I have therefore recommended acceptance
for this version of the manuscript, and I ask that Hickerson et al. provide a
copy of the final version to Jamie Oaks once they have submitted it, so
that he and his co-authors can consider whether they wish to respond.
\end{quote}

In our response, we use simulated and biological data to evaluate the empirical
Bayesian method proposed by Hickerson et al.\ (2014) as a solution to the
problems revealed by Oaks et al.\ (2013) regarding the power and accuracy of the
comparative phylogeographical method implemented in the software
\texttt{msBayes}.
We demonstrate their approach does not mitigate the method's bias toward
supporting models with less parameter space, but rather allows it to manifest
in strong preference for models that often exclude from consideration the true
values of the model's parameters.
Our results also strongly suggest the underlying statistical cause of the
method's tendency to support simple models.
More generally, we discuss the theoretical and practical issues of empirical
Bayesian approaches to model choice; these issues are important for empiricists
applying methods of Bayesian model selection to carefully consider.
Lastly, we discuss ways to mitigate the biases of the \texttt{msBayes} method
while avoiding negative side effects of empirically informed uniform priors.

We performed the work described in the manuscript following the principles of
Open Notebook Science; using version-control software, we make progress in all
aspects of the work freely and publicly available in real-time at
\href{https://github.com/joaks1/msbayes-experiments}{\url{https://github.com/joaks1/msbayes-experiments}}.
All information necessary to reproduce my results is provided there.

Thank you for your consideration, and I look forward to your correspondence
regarding this manuscript.

\addtolength{\medskipamount}{-5pt}
\closing{Sincerely,}
\end{letter}

\end{document}