### 2014-06-01 manuscripts/dpp-msbayes/dpp-msbayes-acknowledgments.tex

\change{
I am greatful to Melissa Callahan, Mark Holder, Emily McTavish, Daniel Money,
Jordan Koch, Adam Leach\'{e}, Peter Foster, two anonymous reviewers, and
Christian Robert and his blog
(\href{http://xianblog.wordpress.com/}{\url{http://xianblog.wordpress.com/}})
for insightful comments that greatly improved this work.
}
I thank the National Science Foundation for supporting this work (DEB
1011423 and DBI 1308885).
This work was also supported by the University of Kansas (KU) Office of Graduate
Studies, Society of Systematic Biologists, Sigma Xi Scientific Research
Society, KU Department of Ecology and Evolutionary Biology, and the KU
Biodiversity Institute.
I also thank Mark Holder, the KU Information and Telecommunication Technology
Center, KU Computing Center, and the iPlant Collaborative for the computational
support necessary to conduct the analyses presented herein.