### 2014-06-16 manuscripts/dpp-msbayes/dpp-msbayes-si-tables.tex
Making some edits to dpp ms in proofing stage.
Adding Table S1 to supporting information that give an example of the
divergence model sample space when Y=4 taxa.

Also, making some edits to the method section where the sample space of
ordered/unordered divergence models is introduced. Making the language
in this section more exact and introducing my loose use of "ordered" and
"unordered" subsequent to this section.

%:TABLE-model-sample-spaces
\begin{table}%[htbp]
    % \small
    % \addtolength{\tabcolsep}{-0.9mm}
    \centering
    \rowcolors{2}{}{myGray} \mcrowcolors
    \captionsetup{name=Table S, labelformat=noSpace}
    \caption{An example showing the number of divergence events (\divTimeNum)
        and the associated sample space of the unordered divergence models
        (integer partitions of \npairs{} pairs) and ordered divergence models
        (partitions of \npairs{} pairs) for $\npairs{} = 4$ pairs of
        populations.}
    \begin{tabular}{ c | l | l }
        \divTimeNum & Unordered divergence models & Ordered divergence models \\
        \hline
        1 & $4$          & 1111                                     \\
        2 & $3+1$; $2+2$ & 1112, 1121, 1211, 2111, 1122, 1212, 1221 \\
        3 & $2+1+1$      & 1123, 1213, 1231, 1223, 1232, 1233       \\
        4 & $1+1+1+1$    & 1234                                     \\
    \end{tabular}
    \label{tabSampleSpace}
\end{table}