### 2014-08-07 manuscripts/reply/cover-letter/cover-letter.tex
More work on reply ms and response to reviews.

\documentclass[letterpaper]{letter}
\usepackage{anysize}
\marginsize{1in}{1in}{0in}{0.5in}
\usepackage{setspace}
\usepackage{url}
\usepackage{verbatim}
\usepackage{parskip}
\usepackage{hyperref}
\hypersetup{pdfborder={0 0 0}, colorlinks=true, urlcolor=blue}

\newcommand{\ignore}[1]{}
\newcommand{\super}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\sub}[1]{\ensuremath{_{\textrm{#1}}}}
\newcommand{\dC}{\ensuremath{^\circ{\textrm{C}}}}

\makeatletter
\let\@texttop\relax
\makeatother

\newenvironment{myEnumerate}{
  \begin{enumerate}
    \setlength{\itemsep}{1pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}}
  {\end{enumerate}}

\newenvironment{myItemize}{
  \begin{itemize}
    \setlength{\itemsep}{1pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}}
  {\end{itemize}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\signature{Jamie Oaks \\
                 Postdoctoral Fellow \\ 
                 Department of Biology \\
                 University of Washington \\ 
                 Box 351800 \\ 
                 Seattle, WA 98195 \\ 
                 USA \\
                 \href{mailto:joaks1@uw.edu}{\texttt{joaks1@uw.edu}} \\ 
                 \href{http://www.phyletica.com}{\texttt{www.phyletica.com}}}
%\address{Biodiversity Institute \\ University of Kansas \\ Dyche Hall, 1345 Jayhawk Blvd \\ Lawrence, KS 66045}
\begin{document}
\begin{letter}{The Society for the Study of Evolution \\
                     PO Box 299 \\
                     St. Louis, MO 63166-0299}
\opening{Dear Editor:}
% \raggedright
I am re-submitting a manuscript (previous submission ID 14-0137) entitled
``Implications of uniformly distributed, empirically informed priors for
phylogeographical model selection:
A reply to Hickerson et al.'' to be considered for publication in the
\emph{International Journal of Organic Evolution}.
This manuscript is a response to a paper by Hickerson et al.\ (2014,
\emph{Evolution} 68:284--294), which was a response to a paper published last
year by my co-authors and me (Oaks et al.\ 2013, \emph{Evolution}
67:991--1010).
Former Associate Editor Dr.\ Laura Kubatko invited us to write this response.

In the manuscript, we highlight alternative hypotheses for why the popular
comparative phylogeographcial method, \texttt{msBayes}, often incorrectly
infers simultaneous divergence across co-distributed species.
Using analyses of simulated and empirical data, we test the distinct
predictions of these hypotheses and find the cause of the bias is not
approximation error, but rather strongly weighted marginal likelihoods of
divergence models under uniform priors on divergence times.
We also evaluate the empirical Bayesian method proposed by Hickerson et al.\
(2014) as a solution to the problems revealed by Oaks et al.\ (2013) regarding
the power and accuracy of the method.
We demonstrate their approach does not mitigate the method's bias toward
supporting models with less parameter space, but rather allows it to manifest
in strong preference for models that often exclude from consideration the true
values of the model's parameters.
More generally, we discuss the theoretical and practical issues of empirical
Bayesian approaches to model choice; these issues are important for empiricists
applying methods of Bayesian model selection to carefully consider.
We highlight ways to mitigate the biases of the \texttt{msBayes} method while
avoiding negative side effects of empirically informed uniform priors.
Overall, this paper highlights the importance of choosing prior assumptions
about parameters in Bayesian selection of models of evolution; if our prior
assumptions place too much weight in unlikely regions of parameter space, the
true posterior will often support the wrong model of evolutionary history.

We performed the work described in the manuscript following the principles of
Open Notebook Science; using version-control software, we make progress in all
aspects of the work freely and publicly available in real-time at
\href{https://github.com/joaks1/msbayes-experiments}{\url{https://github.com/joaks1/msbayes-experiments}}.
All information necessary to reproduce our results is provided there.

The previous submission of this manuscript was handled by Associate Editor Dr.\
Luke Harmon who obtained three anonymous reviews.
We have incorporated the helpful revisions recommended by Dr.\ Harmon and the
reviewers.
My point-by-point response to their comments follows this letter.
Thank you for your consideration, and I look forward to your correspondence
regarding this manuscript.

\addtolength{\medskipamount}{-5pt}
\closing{Sincerely,}
\end{letter}

\end{document}