### 2014-08-07 manuscripts/reply/reply-abstract.tex
More tweaking of abstract for reply ms.

Establishing that a set of population splitting events occurred at the same
time can be a potentially persuasive argument that a common process affected
the populations.
Recently, \citet{Oaks2012} assessed the ability of an approximate-Bayesian
model choice method (\msb) to estimate such a pattern of simultaneous
divergence across taxa, to which \citet{Hickerson2013} responded.
Both papers agree that the primary inference enabled by the method is very
sensitive to prior assumptions and often erroneously supports shared
divergences across taxa when prior uncertainty about divergence times
is represented by a uniform distribution.
However, the papers differ about the best explanation and solution for this
problem.
\citet{Hickerson2013} suggested numerical approximation error causes \msb
analyses to be biased toward models of clustered divergences because the the
method's rejection algorithm is unable to adequately sample the parameter space
of richer models within reasonable computational limits when using broad
uniform priors on divergence times (Hypothesis~\ref{hypError}).
As a potential solution, they proposed a model-averaging approach that uses
narrow, empirically informed uniform priors.
\citet{Oaks2012} suggested the method's behavior was caused by the strong
weight of uniformly distributed priors on divergence times leading to smaller
marginal likelihoods (and thus smaller posterior probabilities) of models with
more divergence time parameters (Hypothesis~\ref{hypML}); they proposed
alternative prior probability distributions to avoid such strongly weighted
posteriors.
% \citet{Oaks2012} suggested the method's behavior was due to the sensitivity of
% marginal likelihoods to the uniformly distributed priors on divergence times
% (Hypothesis~\ref{hypML}), and proposed alternative prior probability
% distributions to mitigate this sensitivity.
Here, we use analyses of simulated and empirical data to show that the tendency
of \msb analyses to support models of shared divergences is primarily due to
Hypothesis~\ref{hypML}, whereas Hypothesis~\ref{hypError} is an untenable
explanation for the bias.
I.e., if our prior assumptions place too much weight in unlikely regions of
parameter space such that the true posterior supports the wrong model of
evolutionary history, no amount of computation can rescue our inference. 
We also demonstrate the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} does not mitigate the method's tendency to erroneously
support models of highly clustered divergences.
Furthermore, we show this approach is dangerous in the sense that the
empirically-derived uniform priors often exclude from consideration the true
values of the divergence-time parameters.
Fortunately, as predicted by fundamental principles of Bayesian model choice,
more flexible distributions that accommodate prior uncertainty about parameters
without placing excessive weight in vast regions of parameter space with low
likelihood increase the method's robustness and power to detect temporal
variation in divergences \citep{Oaks2014dpp}.