### 2015-02-18 slides/bordeaux/slides.tex
Adding Bordeaux slides.

\input{../utils/preamble.tex}
\input{../../manuscripts/utils/macros.tex}
\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
\bibliography{../../manuscripts/bib/references}

% \title[Estimating shared history]{An Improved Approximate-Bayesian Method for Estimating Shared
% Evolutionary History}

\title[Inferring community history]{Inferring Community-scale Diversification
    from Genomic Data}

\author[J.\ Oaks]{
    Jamie R.\ Oaks\inst{1}
}
\institute[University of Washington]{
    \inst{1}%
        Department of Biology, University of Washington
}

% \date{\today}
\date{February 18, 2015}

\begin{document}
\maketitle

{
\setbeamercolor{background canvas}{bg=black}
\setbeamercolor{whitetext}{fg=white}
\begin{frame}
    \frametitle{Processes of diversification}
    \begin{columns}[c]
        \column{.5\textwidth}
        {\usebeamercolor[fg]{whitetext}
        \begin{itemize}
            \item<2-> Community-scale processes are important in
                biodiversification and community assembly
            \item<3-> Accounting for such processes will better our
                understanding of evolutionary history and
                community structure
            \item<4-> We need methods for inferring evolutionary patterns
                predicted by historical events from contemporary populations
        \end{itemize}
        }
        \column{.5\textwidth}
        \begin{figure}
            \begin{center}
            \includegraphics[width=\textwidth]{../images/earth-image.jpg}
            \end{center}
        \end{figure}
    \end{columns}
\end{frame}
}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-present.png}}
\begin{frame}
    % \frametitle{Climate-driven diversification}

\end{frame}
}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
\begin{frame}
    % \frametitle{Climate-driven diversification}    
    \begin{columns}
        \column{0.6\textwidth}

        \vspace{6.5cm}

        \ \\

        \column{0.4\textwidth}

        \vspace{-2cm}

        \begin{uncoverenv}<2->
        \colorbox{white}{
            \begin{minipage}[t]{1.0\textwidth}
                \raggedright
                \textbf{Did repeated fragmentation of islands during
                    inter-glacial rises in sea level promote diversification?}
            \end{minipage}
        }
        \end{uncoverenv}
    \end{columns}
\end{frame}
}

\begin{frame}
    \frametitle{Climate-driven diversification}
\begin{columns}[c]
    \column{.5\textwidth}
        \vspace{-1cm}
        \begin{onlyenv}<1>
        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
        \centerline{
        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/rafe.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/rob.jpg}}
        \centerline{
        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/charles.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/cam.jpg}}
        \centerline{
        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/jake.jpg}}
        \centerline{
        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/allie.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/Luke.jpg}
        \hspace{0.3mm}
        \includegraphics<1>[height=2cm]{/home/jamie/Dropbox/field-photos/people/anthony.png}}
        \end{minipage}
        \end{onlyenv}

        \begin{onlyenv}<2->
        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
        \hspace{0.3mm}
        \includegraphics<2->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
        \end{minipage}
        \end{onlyenv}

    \column{.5\textwidth}
        \vspace{-1cm}
        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
        \includegraphics<1-2>[width=\textwidth]{../images/maps/Philippines.png}
        \includegraphics<3>[width=\textwidth]{../images/maps/Philippines-negros_panay.png}
        % \begin{onlyenv}<4->
        %     \begin{itemize}
        %         \item Strong support for simultaneous divergence of all 22 taxon pairs

        %         \item Posterior probability $>$ 0.96 \\

        %         \item $\sim$100,000--250,000 years ago
        %     \end{itemize}
        % \end{onlyenv}
        \end{minipage}
\end{columns}
\end{frame}

\begin{frame}[t]
    \frametitle<1-4>{Climate-driven diversification: Predictions}
    \frametitle<5->{Divergence-model choice}

    \centering{
    \includegraphics<1>[height=7.5cm]{../images/dmc-cartoon-islands.pdf}
    \includegraphics<2>[height=7.5cm]{../images/dmc-cartoon-taxa.pdf}
    \includegraphics<3>[height=7.5cm]{../images/dmc-cartoon-sea-level.pdf}
    \includegraphics<4>[height=7.5cm]{../images/dmc-cartoon-shared-no-labels.pdf}
    \includegraphics<5>[height=7.5cm]{../images/dmc-cartoon-shared.pdf}
    \includegraphics<6>[height=7.5cm]{../images/dmc-cartoon-2-1.pdf}
    \includegraphics<7>[height=7.5cm]{../images/dmc-cartoon-2-2.pdf}
    \includegraphics<8>[height=7.5cm]{../images/dmc-cartoon-2-3.pdf}
    \includegraphics<9>[height=7.5cm]{../images/dmc-cartoon-general.pdf}
    }
\end{frame}

\begin{frame}[t]
    \frametitle{Divergence-model choice}

    \vspace{-7mm}

    \begin{minipage}[t][0.1\textheight][c]{1.1\linewidth}
        \begin{adjustwidth}{-0.5em}{}
            \begin{tabular}{ p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} p{2.1cm} }
                $m_1$ & $m_2$ & $m_3$ & $m_4$ & $m_5$ \\
            \end{tabular}
        \end{adjustwidth}
    \end{minipage}

    \vspace{-2mm}

    \begin{minipage}[t][0.35\textheight][c]{1.1\linewidth}
        \begin{adjustwidth}{-2.3em}{}
        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-shared.pdf}
        \hspace{0.2mm}
        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-1.pdf}
        \hspace{0.2mm}
        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-2.pdf}
        \hspace{0.2mm}
        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-2-3.pdf}
        \hspace{0.2mm}
        \includegraphics[height=2.9cm]{../images/dmc-cartoon-no-islands-general.pdf}
        \end{adjustwidth}
    \end{minipage}

    \vspace{3.5mm}

    \begin{minipage}[c][0.35\textheight][c]{\linewidth}
        \begin{uncoverenv}<2->
            \begin{center}
            We want to infer \textcolor{blue}{\divModel{}} and
            \textcolor{blue}{\divTimeMapVector} given DNA sequence
            alignments
            \textcolor{blue}{\alignmentVector}
        \end{center}
        \end{uncoverenv}

        \vspace{4mm}

        \begin{uncoverenv}<3->
            \begin{displaybox}[0.85\linewidth]
                \begin{minipage}[c][0.12\textheight][c]{\linewidth}

                \[
                    p(\divTimeMapVector,
                      \divModel{i},
                      \allParameters{}
                      \given \alignmentVector)
                      =
                    \frac{
                        p(\alignmentVector \given
                          \divTimeMapVector,
                          \allParameters{},
                          \divModel{i})
                        p(\divTimeMapVector,
                          \allParameters{}
                          \given \divModel{i})
                        p(\divModel{i})
                        }{p(\alignmentVector)}
                \]
                \vspace{-1mm}
                \end{minipage}
            \end{displaybox}
        \end{uncoverenv}
    \end{minipage}
\end{frame}

\begin{frame}
    \frametitle{New method: \dppmsbayes}
    \begin{itemize}
        % \item<1-> Reparameterized the model implemented in \msb
        \item<1-> Dirichlet-process prior (DPP) over all possible divergence
            models
        \item<2-> Flexible priors on continuous parameters to avoid strongly
            weighted posteriors
        \item<3-> Multi-processing interface to improve performance for
            accommodating genomic datasets
        % \item<4-> Uniform prior over divergence models
    \end{itemize}
    \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\begin{frame}
    \frametitle{\dppmsbayes: Performance}
    \begin{itemize}
        \item<1-> Extensive computer simulations to compare performance to previous
            state-of-the-art methods (\msb).
        \item<2-> New method for estimating shared evolutionary history shows
            improved:
            \begin{enumerate}
                \item<2-> Improved model-choice accuracy 
                \item<2-> More robustness to model violations
                \item<2-> Greater power to detect variation in divergence times
                \item<2-> Faster!
            \end{enumerate}
    \end{itemize}

    \uncover<3->{
    \textbf{\dppmsbayes allows biologists to leverage comparative genomic data
        to assess the affects of community-scale processes on biodiversity}
    }
    \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../images/maps/se-asia-120.png}}
\begin{frame}
    % \frametitle{Climate-driven diversification}    
    \begin{columns}
        \column{0.6\textwidth}

        \vspace{6.5cm}

        \ \\

        \column{0.4\textwidth}

        \vspace{-2cm}

        \begin{uncoverenv}<1->
        \colorbox{white}{
            \begin{minipage}[t]{1.0\textwidth}
                \raggedright
                \textbf{Did repeated fragmentation of islands during
                    inter-glacial rises in sea level promote diversification?}
            \end{minipage}
        }
        \end{uncoverenv}
    \end{columns}
\end{frame}
}

\begin{frame}
    \frametitle{Climate-driven diversification}
\begin{columns}[c]
    \column{.5\textwidth}
        \vspace{-1cm}
        \begin{onlyenv}<1->
        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
        \centerline{
        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-negrina-JAEsselstyn.jpg}
        \hspace{0.3mm}
        \includegraphics<1->[height=1.3cm]{../images/photos/crocidura-beatus-DSBalete.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1->[height=1.3cm]{../images/photos/hipposideros-obscurus-MRMDuya.jpg}
        \hspace{0.3mm}
        \includegraphics<1->[height=1.3cm]{../images/photos/haplonycteris-fischeri-JHolden.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-abdictus-rmb.jpg}
        \hspace{0.3mm}
        \includegraphics<1->[height=1.3cm]{../images/photos/sphenomorphus-arborens-rmb.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1->[height=1.3cm]{../images/photos/gekko-mindorensis.jpg}
        \hspace{0.3mm}
        \includegraphics<1->[height=1.3cm]{../images/photos/dendrelaphis-pictus-cds.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-agusanensis.jpg}
        \hspace{0.3mm}
        \includegraphics<1->[height=1.3cm]{../images/photos/cyrt-annulatus-cds.jpg}}
        % \vspace{0.5mm}
        \centerline{
        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-magnus-cds.jpg}
        \hspace{0.3mm}
        \includegraphics<1->[height=1.3cm]{../images/photos/limnonectes-leytensis-rmb.jpg}}
        \end{minipage}
        \end{onlyenv}

    \column{.5\textwidth}
        \vspace{-1cm}
        \begin{minipage}[t][1.0\textheight][c]{\linewidth}
        \includegraphics<1->[width=\textwidth]{../images/maps/Philippines.png}
        \end{minipage}
\end{columns}
\end{frame}

\begin{frame}
    \frametitle{Results}
    \centerline{
    \includegraphics[width=\textwidth]{../../empirical-analyses/plots/philippines-dpp-psi-posterior-prior.pdf}}
    \smallskip
    \centerline{
    \includegraphics[height=2cm]{../images/sea-level-only.pdf}}
    \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\begin{frame}
    \frametitle{Current work: More data}
    \begin{itemize}
        \item<1-> Collecting genomic data from taxa co-distributed across Southeast
            Asia
        \item<2-> Preliminary results for 1000 loci from 5 pairs of \emph{Gekko
                mindorensis} populations
    \end{itemize}

    \begin{center}
        \includegraphics<2->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
    \end{center}
\end{frame}

\begin{frame}
    \frametitle{Current work: More power}
    \centerline{
    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/mtholder.jpg}
    \hspace{0.6mm}
    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
    \hspace{0.6mm}
    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/vladimir.jpg}
    \hspace{0.6mm}
    \includegraphics<1->[height=2cm]{../images/revbayes.png}}

    \begin{itemize}[<+->]
        \item Full-likelihood Bayesian implementation
            \begin{itemize}
                \item Uses all the information in the data
                \item Applicable to deeper timescales
            \end{itemize}
        \item Analytically integrate over gene trees
            \footnote{\tiny\shortfullcite{Bryant2012}}
            \begin{itemize}
                \item Very efficient numerical approximation of posterior 
                \item Applicable to NGS datasets
            \end{itemize}
    \end{itemize}
\end{frame}

\section{Conclusions}

\begin{frame}
    \frametitle{Climate-driven diversification model?}
    \begin{itemize}
        \item<1-> Results consistent with prediction of clustered
            divergences
        \item<2-> Results suggest multiple co-divergences
        \item<3-> However, there is a lot of uncertainty!
    \end{itemize}
    \barefootnote{\shortfullcite{Oaks2014dpp}}
\end{frame}

\section{Teaching}

\begin{frame}
    BIO 180 stuff
\end{frame}

\begin{frame}
    Software carpentry \\

    Workshops
\end{frame}

\begin{frame}
    \frametitle{Questions?}    
    \begin{center}
        {
        \Large
        \href{mailto:joaks1@gmail.com}{\tt joaks1@gmail.com}
        }
    \end{center}
\end{frame}

\begin{frame}
    \frametitle{Future directions---Computational}
    \begin{itemize}
        \item<1-> Full-phylogenetic framework
    \end{itemize}

    \begin{center}
        \includegraphics<1>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
        \includegraphics<2>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
    \end{center}
\end{frame}

\begin{frame}
    \centerline{
    \includegraphics<1>[height=8cm]{../images/bd-cartoon-no-lines.pdf}
    \includegraphics<2>[height=8cm]{../images/bd-cartoon-shared.pdf}
    }
\end{frame}

\begin{frame}
\begin{columns}
    \column{.5\textwidth}
    \begin{minipage}[c]{\columnwidth}
    \begin{itemize}
        \item Blah
    \end{itemize}
    \end{minipage}

    \column{.6\textwidth}
    \begin{minipage}[t][\textheight][c]{\linewidth}
    \includegraphics<1>[width=0.75\columnwidth]{../images/bd-cartoon-shared-space.pdf}
    \includegraphics<2>[width=0.75\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
    \end{minipage}
\end{columns}
\end{frame}

\end{document}