### 2015-02-18 slides/bordeaux/slides-2.tex
Working on bordeaux slides

\input{../utils/preamble.tex}
\input{../../manuscripts/utils/macros.tex}
\newcommand{\allParameters}[1]{\ensuremath{\theta_{#1}}\xspace}
\bibliography{../../manuscripts/bib/references}

% \title[Estimating shared history]{An Improved Approximate-Bayesian Method for Estimating Shared
% Evolutionary History}

\title[Future directions]{Future Directions}

\author[J.\ Oaks]{
    Jamie R.\ Oaks\inst{1}
}
\institute[University of Washington]{
    \inst{1}%
        Department of Biology, University of Washington
}

% \date{\today}
\date{February 18, 2015}

\begin{document}

\begin{frame}[t]
    \frametitle{New methods for inferring community history}
    \vspace{-5mm}
    \begin{minipage}[t][0.1\textheight][c]{1\linewidth}
    \begin{itemize}
        \item<2-> Incorporate comparative models of community-scale
            co-diversification into a full-phylogenetic framework
    \end{itemize}
    \end{minipage}

    \vspace{-5mm}
    \begin{minipage}[t][0.9\textheight][c]{1\linewidth}
    \begin{center}
        \includegraphics<3>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
        \includegraphics<4>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
    \end{center}
    \end{minipage}
\end{frame}

\begin{frame}
    \begin{columns}
        \column{0.4\textwidth}
        \begin{minipage}[c]{\columnwidth}
        \begin{adjustwidth}{-2.3em}{}
        \begin{onlyenv}<1-6>
        \begin{itemize}
            \item<2-> Current phylogenetic methods assume speciation and extinction
                occur independently of one another
            \item<3-> We know this assumption is frequently violated
            \item<4-> I will generalize a birth-death model to accommodate
                ``shared'' events
            \item<6> Infer events jointly across multiple trees
        \end{itemize}
        \end{onlyenv}
        \begin{onlyenv}<7->
        \only<7->{\hspace{0.5cm}\textbf{\large Benefits:}\\}
        \begin{itemize}
            \item<8-> Improve phylogenetic inference and divergence-time estimation
            \item<9-> Provide a general framework for inferring the evolutionary
                history of whole communities
            \item<10-> Possible epidemiological applications
        \end{itemize}
        \end{onlyenv}
        \end{adjustwidth}
        \end{minipage}

        \column{0.6\textwidth}
        \begin{minipage}[t][\textheight][c]{\linewidth}
        \centerline{
        \includegraphics<1-3>[width=\columnwidth]{../images/bd-cartoon-no-lines.pdf}
        \includegraphics<4>[width=\columnwidth]{../images/bd-cartoon-shared.pdf}
        \includegraphics<5>[width=0.75\columnwidth]{../images/bd-cartoon-shared-space.pdf}
        \includegraphics<6->[width=0.75\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
        }
        \end{minipage}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Relevance to LabEx COTE}
    \begin{itemize}
        \item<2-> Use novel comparative method to estimate the history of floral
            and faunal communities across Europe
        \item<3-> Elucidate the history of communities and their response to past
            environmental climate change
        \item<4-> This will provide context for predicting future change
        \item<5-> Collaboration with BIOGECO partners
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Relevance to LabEx COTE}
    \begin{itemize}[<+->]
        \item Generalized birth-death process will be relevant to
            modeling epidemiological dynamics
        \item Relevant to the evolution of grapevine pathogens
        \item Collaboration with SV partners to investigate disease dynamics
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Strategies for integration}
    \begin{itemize}
        \item<2-> In order to predict the response of ecosystems to environmental
            change, we need a general model of ecosystem processes
        \item<3-> This requires integration across LabEx COTE partners
        \item<3-> Organize a series ``hackathons'' to stimulate synthesis among the
            biologists, engineers, theorists, and modelers across LabEx COTE
        \item<4-> Each hackathon would focus on integrating concepts and existing
            models between particular ecosystems and/or ecosystem processes
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Teaching prospects}
    \begin{itemize}[<+->]
        \item Teach courses at the nexus of evolution, ecology, statistics,
            and computer science
        \begin{itemize}
            \item Computational biology
            \item Advance statistics in biology (likelihood and Bayesian
                approaches)
            \item Advanced methods in evolutionary genetics and
                molecular ecology
        \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Questions?}    
    \begin{columns}[c]
        \column{.5\textwidth}
        \begin{center}
            {
            \Large
            \href{mailto:joaks1@uw.edu}{\tt joaks1@uw.edu}
            }
        \end{center}
        \column{.5\textwidth}
            \begin{figure}
                \begin{center}
                \includegraphics[width=\textwidth]{../images/darwin-tol-copyright-boris-kulikov-2007.jpg}
                \caption{\tiny \copyright~2007 Boris Kulikov \href{http://boris-kulikov.blogspot.com/}{boris-kulikov.blogspot.com}}
                \end{center}
            \end{figure}
    \end{columns}
\end{frame}

\end{document}