### 2015-03-03 slides/auburn/intro-mascot-tree.tex
Working on slides for auburn.

\begin{frame}
    \begin{center}
        % \transwipe<2>[direction=0, duration=3]
        % \includegraphics<2>[width=\textwidth]{../images/mascot-tree.png}
        \includegraphics[width=\textwidth]{../images/mascot-tree.png}

    \end{center}
\end{frame}