### 2015-03-04 slides/auburn/intro-mascot-tree.tex
Working on slides.

\begin{frame}
    \begin{center}
        % \transwipe<2>[direction=0, duration=3]
        % \includegraphics<2>[width=\textwidth]{../images/mascot-tree.png}
        \includegraphics<1>[width=\textwidth]{../mascot-tree/mascot-tree-shared.pdf}
        \includegraphics<2->[width=\textwidth]{../mascot-tree/mascot-tree-shared-events.pdf}

    \end{center}
\end{frame}