### 2015-03-04 slides/auburn/abc.tex
Working on slides.

\begin{frame}
\newcommand{\matrixonebottom}{6.0}
\newcommand{\matrixtwobottom}{3.0}
\newcommand{\matrixthreebottom}{0.0}
\newcommand{\matrixspace}{0.285}
% \begin{picture}(300,200)(0,0)
    % \put(0,-34){
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2.2cm,on grid,auto]
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*9) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*8) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*7) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*6) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*5) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1a.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*4) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*3) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*2) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*1) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};
	\node [visible on=<1->] at (0,\matrixonebottom+\matrixspace*0) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-1b.pdf}}};

	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*9) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*8) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*7) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*6) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*5) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2a.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*4) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*3) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*2) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*1) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};
	\node [visible on=<1->] at (0,\matrixtwobottom+\matrixspace*0) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-2b.pdf}}};

	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*9) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*8) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*7) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*6) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*5) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3a.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*4) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*3) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*2) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*1) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
	\node [visible on=<1->] at (0,\matrixthreebottom+\matrixspace*0) (l1a) {\pgftext{\includegraphics[height=3mm]{../images/lizard-3b.pdf}}};
	\node [visible on=<2->] at (2.6,1.3) (m3) {\pgftext{\includegraphics[height=28mm]{../images/colormat1.pdf}}};
	\node [visible on=<2->] at (2.6,4.3) (m2) {\pgftext{\includegraphics[height=28mm]{../images/colormat2.pdf}}};
	\node [visible on=<2->] at (2.6,7.3) (m1) {\pgftext{\includegraphics[height=28mm]{../images/colormat3.pdf}}};

    \node[visible on=<3->] at (8, 7.3) (s1) {$S_1$};
    \node[visible on=<3->] at (8, 4.3) (s2) {$S_2$};
    \node[visible on=<3->] at (8, 1.3) (s3) {$S_3$};

    \path[->] (5.2,7.3) edge [visible on=<3->,very thick, bend left] (s1);
    \path[->] (5.2,4.3) edge [visible on=<3->,very thick, bend left] (s2);
    \path[->] (5.2,1.3) edge [visible on=<3->,very thick, bend left] (s3);

\end{tikzpicture}
\end{center}
    % }
% \end{picture}
\end{frame}

\begin{frame}[t]
    % \frametitle{Easy as ABC}
    \begin{center}
    \begin{tikzpicture}[shorten >=1pt,node distance=2.2cm,on grid,auto]
        \node[] at (0,0) {\includegraphics<1>[width=\textwidth]{../images/rejection-sampling-observed.pdf}};
        \node[] at (0,0) {\includegraphics<2>[width=\textwidth]{../images/rejection-sampling-tolerance.pdf}};
        \node[] at (0,0) {\includegraphics<3>[width=\textwidth]{../images/rejection-sampling-10.pdf}};
        \node[] at (0,0) {\includegraphics<4>[width=\textwidth]{../images/rejection-sampling-100.pdf}};
        \node[] at (0,0) {\includegraphics<5>[width=\textwidth]{../images/rejection-sampling-200.pdf}};
        \node[] at (0,0) {\includegraphics<6>[width=\textwidth]{../images/rejection-sampling-500.pdf}};
        \node[] at (0,0) {\includegraphics<7>[width=\textwidth]{../images/rejection-sampling-1000.pdf}};
        \node[] at (0,0) {\includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}};
        \node[] at (0,0) {\includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}};
        \node[] at (0,0) {\includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}};
        \node[visible on=<1->] at (-2.9, -3.7) (s1) {$S_1$};
        \node[visible on=<1->] at (3.4, -2.9) (s2) {$S_2$};
        \node[visible on=<1->] at (5.3, 0.7) (s3) {$S_3$};
    \end{tikzpicture}
    \end{center}
    % \centerline{
    % \includegraphics<1>[width=\textwidth]{../images/rejection-sampling-observed.pdf}
    % \includegraphics<2>[width=\textwidth]{../images/rejection-sampling-tolerance.pdf}
    % \includegraphics<3>[width=\textwidth]{../images/rejection-sampling-10.pdf}
    % \includegraphics<4>[width=\textwidth]{../images/rejection-sampling-100.pdf}
    % \includegraphics<5>[width=\textwidth]{../images/rejection-sampling-200.pdf}
    % \includegraphics<6>[width=\textwidth]{../images/rejection-sampling-500.pdf}
    % \includegraphics<7>[width=\textwidth]{../images/rejection-sampling-1000.pdf}
    % \includegraphics<8>[width=\textwidth]{../images/rejection-sampling-10000.pdf}
    % \includegraphics<9>[width=\textwidth]{../images/rejection-sampling-20000.pdf}
    % \includegraphics<10>[width=\textwidth]{../images/rejection-sampling-20000-post.pdf}
    % }
\end{frame}