### 2015-03-05 slides/auburn/dpp-intro.tex

\begin{frame}[t]
    \frametitle{Sampling divergence models---a novel approach}
    \begin{itemize}[<+->]
        \item The divergence models are ways of assigning our taxa to events
        \item A Dirichlet process prior (DPP) model is a ``natural'' fit
            \begin{itemize}[<+->]
                \item Common nonparametric Bayesian approach to assigning
                    variables to an unknown number of categories
            \end{itemize}
    \end{itemize}

    \begin{uncoverenv}<2->
    \begin{center}
        \includegraphics[height=0.4\textheight]{../images/peter-dirichlet.jpg}
    \end{center}
    \end{uncoverenv}
\end{frame}