### 2015-03-05 slides/auburn/full-likelihood.tex

\begin{frame}
    \frametitle{Current work: More power}
    \centerline{
    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/mtholder.jpg}
    \hspace{0.6mm}
    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/jeet2.jpg}
    \hspace{0.6mm}
    \includegraphics<1->[height=2cm]{/home/jamie/Dropbox/field-photos/people/vladimir.jpg}
    \hspace{0.6mm}
    \includegraphics<1->[height=2cm]{../images/revbayes.png}}

    \begin{itemize}
        \item<1-> Full-likelihood Bayesian implementation
            \begin{itemize}
                \item<2-> Uses all the information in the data
                \item<2-> Applicable to deeper timescales
            \end{itemize}
        \item<3-> Analytically integrate over gene trees
            \footnote{\tiny\shortfullcite{Bryant2012}}
            \begin{itemize}
                \item<4-> Very efficient numerical approximation of posterior 
                \item<4-> Applicable to NGS datasets
            \end{itemize}
    \end{itemize}
\end{frame}