### 2015-03-05 slides/auburn/more-data.tex

\begin{frame}[t]
    \frametitle{Current work: More data}
    \begin{minipage}[t][0.15\textheight][c]{\linewidth}
    \begin{itemize}
        \item<1-> Collecting genomic data from taxa co-distributed across
            Southeast Asian Islands and Mainland
        \item<2-> Preliminary results for 1000 loci from 5 pairs of \emph{Gekko
                mindorensis} populations
    \end{itemize}
    \end{minipage}

    \begin{center}
        \includegraphics<2->[height=5.5cm]{../images/number-of-divs-gekko-mindorensis.pdf}
    \end{center}
\end{frame}