### 2015-03-05 slides/auburn/sampling-models.tex

\begin{frame}[t]
    \frametitle{Sampling divergence models---previous approach}
    \begin{itemize}
        \item<2-> Place a discrete uniform prior on the
            \textit{\textbf{number}} of events
    \end{itemize}

    \begin{uncoverenv}<2->
    \begin{columns}
        \column{0.5\textwidth}
            \centerline{
                \includegraphics<2->[width=1.0\textwidth]{../images/number-of-div-models-10.pdf}}
        \column{0.5\textwidth}
            \centerline{
                \includegraphics<3->[width=1.0\textwidth]{../images/prob-of-div-models-10.pdf}}
    \end{columns}
    \end{uncoverenv}

    \begin{itemize}
        \item<4-> This can create a strong bias toward models of co-divergence
    \end{itemize}

    \begin{uncoverenv}<5->
    \begin{block}{\it Potential solution:}
        Place flexible prior directly on the sample space of divergence models
    \end{block}
    \end{uncoverenv}
\end{frame}