### 2015-03-07 slides/auburn/dpp-intro.tex
Working on slides.

\begin{frame}[t]
    \frametitle{Sampling divergence models---a novel approach}
    \begin{itemize}[<+->]
        \item The divergence models are ways of assigning our taxa to events
        \item A Dirichlet process prior (DPP) model is convenient and flexible
            solution
            \begin{itemize}[<+->]
                \item Common Bayesian approach to assigning variables to an
                    unknown number of categories
                \item Controlled by ``concentration'' parameter: $\alpha$
            \end{itemize}
    \end{itemize}

    \begin{uncoverenv}<2->
    \begin{figure}
    \begin{center}
        \includegraphics[height=0.4\textheight]{../images/peter-dirichlet.jpg}
        \caption{Peter Dirichlet}
    \end{center}
    \end{figure}
    \end{uncoverenv}
\end{frame}