### 2015-03-07 slides/auburn/full-phylo.tex

\begin{frame}[t]
    \frametitle{Next step: A general framework}
    \vspace{-5mm}
    \begin{minipage}[t][0.1\textheight][c]{1\linewidth}
    \begin{itemize}
        \item<2-> Incorporate comparative models of community-scale
            co-diversification into a full-phylogenetic framework
    \end{itemize}
    \end{minipage}

    \vspace{-5mm}
    \begin{minipage}[t][0.9\textheight][c]{1\linewidth}
    \begin{center}
        \includegraphics<3>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-shared-middle-wide.pdf}
        \includegraphics<4>[height=6.5cm]{../images/dmc-cartoon-no-islands-pop-phylogeny.pdf}
    \end{center}
    \end{minipage}
\end{frame}

\begin{frame}
    \frametitle{Next step: A general framework}

    \vspace{-5mm}
    \begin{columns}
        \column{0.4\textwidth}
        \begin{minipage}[c]{\columnwidth}
        \begin{adjustwidth}{-2.3em}{}
        \begin{onlyenv}<1-4>
        \begin{itemize}
            \item<1-> Relax the assumption of independence of events across the
                tree
            \item<2-> I will generalize a birth-death model to accommodate
                shared events
            \item<4-> Jointly infer co-diversification across multiple trees
        \end{itemize}
        \end{onlyenv}
        \begin{onlyenv}<5->
        \only<5->{\hspace{0.5cm}\textbf{\large Benefits:}\\}
        \begin{itemize}
            \item<6-> Improve phylogenetic inference and divergence-time estimation
            \item<7-> Provide a general framework for inferring shared
                evolutionary history across communities of species
            \item<8-> Epidemiological applications
        \end{itemize}
        \end{onlyenv}
        \end{adjustwidth}
        \end{minipage}

        \column{0.6\textwidth}
        \begin{minipage}[t][\textheight][c]{\linewidth}
        \centerline{
        \includegraphics<1>[width=\columnwidth]{../images/bd-cartoon-no-lines.pdf}
        \includegraphics<2>[width=\columnwidth]{../images/bd-cartoon-shared.pdf}
        \includegraphics<3>[width=0.7\columnwidth]{../images/bd-cartoon-shared-space.pdf}
        \includegraphics<4->[width=0.7\columnwidth]{../images/bd-cartoon-2-trees-shared.pdf}
        }
        \end{minipage}
    \end{columns}
\end{frame}